MSHS COVID-19 Vaccination Summary

### 2022-04-18 MSH Admin Dose Summary.Rmd
Simplify logic for determining adult vs pediatric vaccine.

sched_to_date <- sched_to_date %>%
  # rename(IsPtEmployee = `Is the patient a Mount Sinai employee`) %>%
  mutate(
    # Update timezones to EST if imported as UTC
    `Appt Time` = with_tz(`Appt Time`, tzone = "EST"),
    # Determine whether appointment is for dose 1, 2, or 3 based on scheduled visit type
    Dose = ifelse(str_detect(Type, "DOSE 1"), "Dose 1",
                  ifelse(str_detect(Type, "DOSE 2"), "Dose 2",
                         ifelse(str_detect(Type, "(DOSE 3)|(BOOSTER)"),
                                "Dose 3/Boosters", NA))),
    # Determine appointment date, year, month, etc.
    ApptDate = date(Date),
    ApptYear = year(Date),
    ApptMonth = month(Date),
    ApptDay = day(Date),
    ApptWeek = week(Date),
    # Clean up department if it is duplicated in that column
    Department = ifelse(str_detect(Department, ","),
                        substr(Department, 1,
                               str_locate(Department, ",") - 1),
                        Department),
    # Group arrived, completed, and checked out appts as arrived
    Status = ifelse(`Appt Status` %in% c("Arrived", "Comp", "Checked Out"),
                    "Arr", `Appt Status`),
    # Update appointment status: classify any appts from prior days that are
    # still in scheduled as No Shows
    Status2 = ifelse(ApptDate < (today) & Status == "Sch", "No Show", 
                     # Correct any appts erroneously marked as arrived early
                     ifelse(ApptDate >= today & Status == "Arr", "Sch",
                            Status)),
    # # Determine whether the vaccine is an adult or pediatric dose
    # VaxType = ifelse(Department %in% c("1468 MADISON PEDIATRIC VACCINE POD",
    #                                    "1468 MADISON HOSP PEDS GEN") |
    #                    `Provider/Resource` %in% c("DOSE 1 PEDIATRIC [1324684]",
    #                                               "DOSE 2 PEDIATRIC [1324685]") |
    #                    Department %in% peds_school_practice_dept,
    #                  "Peds", "Adult"),
    # # Determine manufacturer based on immunization field and vaccine type fields
    # Mfg = #Keep manufacturer as NA if the appointment hasn't been arrived
    #   ifelse(Status2 != "Arr", NA,
    #          # Any blank immunizations and any peds assumed to be Pfizer
    #          ifelse(is.na(Immunizations) |
    #                   VaxType %in% c("Peds"), "Pfizer",
    #                 # Any immunizations with Moderna in text classified as Moderna
    #                 ifelse(str_detect(Immunizations, "Moderna"), "Moderna",
    #                        # Any visiting docs or Johnson and Johnson
    #                        # immunizations classified as J&J
    #                        ifelse(str_detect(Department, "VISITING DOCS") |
    #                                 str_detect(Immunizations,
    #                                            "Johnson and Johnson"),
    #                               "J&J", "Pfizer")))),
    # Determine week number of year
    WeekNum = format(ApptDate, "%U"),
    # Determine appointment day of week
    DOW = weekdays(ApptDate),
    # # Determine if patient's zip code is in NYS
    # NYZip = substr(`ZIP Code`, 1, 5) %in% ny_zips$zipcode,
    # Determine if patient is a MSHS employee based on response to question
    # MSHSEmployee =
    #   # First see if patient responded "YES" to question about MSHS employee
    #   str_detect(
    #     str_replace_na(IsPtEmployee, ""),
    #     regex("yes", ignore_case = TRUE)) |
    #   # Next, see if there is a MSHS site listed for patients who were vaccinated
    #   # prior to the addition of the MSHS employee field
    #   str_detect(
    #     str_replace_na(`MOUNT SINAI HEALTH SYSTEM`, " "),
    #     regex("[a-z]", ignore_case = TRUE))
    )

@@

# Determine vaccine type (adult vs peds) administered based on department, provider, and age, as appropriate
sched_to_date <- sched_to_date %>%
  mutate(
    # Create a column for age grouper
    Age_Grouper = ifelse(str_detect(Age, "\\ y.o."), "years",
                        ifelse(str_detect(Age, "\\ m.o."), "months",
                               ifelse(str_detect(Age, "\\ days"), "days",
                                      "unknown"))),
    # Convert string with age to numeric
    Age_Numeric = 
      ifelse(Age_Grouper == "years",
                    as.numeric(str_extract(Age,
                                 # pattern = "(?<=Deceased\\ \\().+(?=\\ y.o.)")),
                                 paste("(?<=Deceased\\ \\()[0-9]+(?=\\ y.o.)",
                                       "[0-9]+(?=\\ y.o.)",
                                       sep = "|"))),
                    ifelse(Age_Grouper == "months",
                           as.numeric(str_extract(Age, "[0-9]+(?=\\ m.o.)")) / 12,
                           ifelse(Age_Grouper == "days",
                                  as.numeric(
                                    str_extract(Age, "[0-9]+(?=\\ days)")) / 365,
                                  ifelse(Age_Grouper == "unknown", 10000,
                                         NA_integer_)))),
    # Determine vaccine type based on department, provider/resource, and age if necessary
    VaxType = ifelse(
      `Vaccine Age Group` %in% "Peds" |
        (`Vaccine Age Group` %in% "Both" &
           # Determine if provider/resource at MSBI is peds
           (`Provider/Resource` %in% c("DOSE 1 PEDIATRIC [1324684]",
                                       "DOSE 2 PEDIATRIC [1324685]") |
              (`Setting Grouper` %in% c("School Based Practice", "MSHS Practice") &
                 Age_Numeric < 12))),
      "Peds", "Adult"),
    # Determine manufacturer based on immunization field and vaccine type fields
    Mfg = #Keep manufacturer as NA if the appointment hasn't been arrived
      ifelse(Status2 != "Arr", NA,
             # Any blank immunizations and any peds assumed to be Pfizer
             ifelse(is.na(Immunizations) |
                      VaxType %in% c("Peds"), "Pfizer",
                    # Any immunizations with Moderna in text classified as Moderna
                    ifelse(str_detect(Immunizations, "Moderna"), "Moderna",
                           # Any visiting docs or Johnson and Johnson
                           # immunizations classified as J&J
                           ifelse(str_detect(Department, "VISITING DOCS") |
                                    str_detect(Immunizations,
                                               "Johnson and Johnson"),
                                  "J&J", "Pfizer")))),
    Site =
      # Roll up any Moderna administered at MSB for YWCA to MSH
      ifelse(Site %in% c("MSB") & Mfg %in% c("Moderna"), "MSH",
             # Manually classify school practices as peds or adolescent for schools administering both
             ifelse(`Setting Grouper` %in% "School Based Practice" &
                      VaxType %in% "Peds", "Peds School Based Practice",
                    ifelse(`Setting Grouper` %in% "School Based Practice" &
                             VaxType %in% "Adult",
                           "Adolescent School Based Practice",
                           Site))),
    # Manually correct any J&J doses erroneously documented as dose 2 and change to dose 1
    Dose = ifelse(Mfg %in% c("J&J") & Dose %in% c("Dose 2"), "Dose 1", Dose),
    `Setting Grouper` = ifelse(`Setting Grouper` %in% "School Based Practice",
                               "MSHS Practice", `Setting Grouper`))

>>>>

sched_to_date <- sched_to_date %>%
  # rename(IsPtEmployee = `Is the patient a Mount Sinai employee`) %>%
  mutate(
    # Update timezones to EST if imported as UTC
    `Appt Time` = with_tz(`Appt Time`, tzone = "EST"),
    # Determine whether appointment is for dose 1, 2, or 3 based on scheduled visit type
    Dose = ifelse(str_detect(Type, "DOSE 1"), "Dose 1",
                  ifelse(str_detect(Type, "DOSE 2"), "Dose 2",
                         ifelse(str_detect(Type, "(DOSE 3)|(BOOSTER)"),
                                "Dose 3/Boosters", NA))),
    # Determine appointment date, year, month, etc.
    ApptDate = date(Date),
    ApptYear = year(Date),
    ApptMonth = month(Date),
    ApptDay = day(Date),
    ApptWeek = week(Date),
    # Clean up department if it is duplicated in that column
    Department = ifelse(str_detect(Department, ","),
                        substr(Department, 1,
                               str_locate(Department, ",") - 1),
                        Department),
    # Group arrived, completed, and checked out appts as arrived
    Status = ifelse(`Appt Status` %in% c("Arrived", "Comp", "Checked Out"),
                    "Arr", `Appt Status`),
    # Update appointment status: classify any appts from prior days that are
    # still in scheduled as No Shows
    Status2 = ifelse(ApptDate < (today) & Status == "Sch", "No Show", 
                     # Correct any appts erroneously marked as arrived early
                     ifelse(ApptDate >= today & Status == "Arr", "Sch",
                            Status)),
    # # Determine whether the vaccine is an adult or pediatric dose
    # VaxType = ifelse(Department %in% c("1468 MADISON PEDIATRIC VACCINE POD",
    #                                    "1468 MADISON HOSP PEDS GEN") |
    #                    `Provider/Resource` %in% c("DOSE 1 PEDIATRIC [1324684]",
    #                                               "DOSE 2 PEDIATRIC [1324685]") |
    #                    Department %in% peds_school_practice_dept,
    #                  "Peds", "Adult"),
    # # Determine manufacturer based on immunization field and vaccine type fields
    # Mfg = #Keep manufacturer as NA if the appointment hasn't been arrived
    #   ifelse(Status2 != "Arr", NA,
    #          # Any blank immunizations and any peds assumed to be Pfizer
    #          ifelse(is.na(Immunizations) |
    #                   VaxType %in% c("Peds"), "Pfizer",
    #                 # Any immunizations with Moderna in text classified as Moderna
    #                 ifelse(str_detect(Immunizations, "Moderna"), "Moderna",
    #                        # Any visiting docs or Johnson and Johnson
    #                        # immunizations classified as J&J
    #                        ifelse(str_detect(Department, "VISITING DOCS") |
    #                                 str_detect(Immunizations,
    #                                            "Johnson and Johnson"),
    #                               "J&J", "Pfizer")))),
    # Determine week number of year
    WeekNum = format(ApptDate, "%U"),
    # Determine appointment day of week
    DOW = weekdays(ApptDate)
    )

@@

# Determine vaccine type (adult vs peds) administered based on department, provider, and age, as appropriate
sched_to_date <- sched_to_date %>%
  mutate(
    # Create a column for age grouper
    Age_Grouper = ifelse(str_detect(Age, "\\ y.o."), "years",
                        ifelse(str_detect(Age, "\\ m.o."), "months",
                               ifelse(str_detect(Age, "\\ days"), "days",
                                      "unknown"))),
    # Convert string with age to numeric
    Age_Numeric = 
      ifelse(Age_Grouper == "years",
                    as.numeric(str_extract(Age,
                                 # pattern = "(?<=Deceased\\ \\().+(?=\\ y.o.)")),
                                 paste("(?<=Deceased\\ \\()[0-9]+(?=\\ y.o.)",
                                       "[0-9]+(?=\\ y.o.)",
                                       sep = "|"))),
                    ifelse(Age_Grouper == "months",
                           as.numeric(str_extract(Age, "[0-9]+(?=\\ m.o.)")) / 12,
                           ifelse(Age_Grouper == "days",
                                  as.numeric(
                                    str_extract(Age, "[0-9]+(?=\\ days)")) / 365,
                                  ifelse(Age_Grouper == "unknown", 10000,
                                         NA_integer_)))),
    # Determine vaccine type based on department, provider/resource, and age if necessary
    VaxType = ifelse(
      `Vaccine Age Group` %in% "Peds" |
        # MSBI Peds POD
        `Provider/Resource` %in% c("DOSE 1 PEDIATRIC [1324684]",
                                       "DOSE 2 PEDIATRIC [1324685]") |
        # Practics that administer adult and pediatric vaccines
        (`Vaccine Age Group` %in% "Both" &
              `Setting Grouper` %in% c("School Based Practice", "MSHS Practice") &
                 Age_Numeric < 12),
      "Peds", "Adult"),
    # Determine manufacturer based on immunization field and vaccine type fields
    Mfg = #Keep manufacturer as NA if the appointment hasn't been arrived
      ifelse(Status2 != "Arr", NA,
             # Any blank immunizations and any peds assumed to be Pfizer
             ifelse(is.na(Immunizations) |
                      VaxType %in% c("Peds"), "Pfizer",
                    # Any immunizations with Moderna in text classified as Moderna
                    ifelse(str_detect(Immunizations, "Moderna"), "Moderna",
                           # Any visiting docs or Johnson and Johnson
                           # immunizations classified as J&J
                           ifelse(str_detect(Department, "VISITING DOCS") |
                                    str_detect(Immunizations,
                                               "Johnson and Johnson"),
                                  "J&J", "Pfizer")))),
    Site =
      # Roll up any Moderna administered at MSB for YWCA to MSH
      ifelse(Site %in% c("MSB") & Mfg %in% c("Moderna"), "MSH",
             # Manually classify school practices as peds or adolescent for schools administering both
             ifelse(`Setting Grouper` %in% "School Based Practice" &
                      VaxType %in% "Peds", "Peds School Based Practice",
                    ifelse(`Setting Grouper` %in% "School Based Practice" &
                             VaxType %in% "Adult",
                           "Adolescent School Based Practice",
                           Site))),
    # Manually correct any J&J doses erroneously documented as dose 2 and change to dose 1
    Dose = ifelse(Mfg %in% c("J&J") & Dose %in% c("Dose 2"), "Dose 1", Dose),
    `Setting Grouper` = ifelse(`Setting Grouper` %in% "School Based Practice",
                               "MSHS Practice", `Setting Grouper`))
